MATH-469 Project Checkin

### 2012-11-10 checkin/project-checkin.tex
checkin

\documentclass[letterpaper,12pt]{article}
\usepackage[left=3cm,right=3cm,top=3cm,bottom=3cm]{geometry}
\usepackage{enumerate}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{float}
\usepackage[usenames]{xcolor}

@@

What follows is an explanation of the initial stages of the algorithm on
64-character genome and three r

@@

\begin{figure}
\includegraphics{out.digraph.pdf}
\end{figure}

\begin{figure}
\includegraphics{out.bidigraph.pdf}
\end{figure}

\begin{figure}
\includegraphics{out.reduced.digraph.pdf}
\end{figure}

\begin{figure}
\includegraphics{out.reduced.bidigraph.pdf}
\end{figure}

\begin{figure}
\includegraphics{out.reduced.collapsed.digraph.pdf}
\end{figure}

\begin{figure}
\includegraphics{out.reduced.collapsed.bidigraph.pdf}
\end{figure}

>>>>

\documentclass[letterpaper,12pt]{article}
\usepackage[left=3cm,right=3cm,top=3cm,bottom=3cm]{geometry}
\usepackage{enumerate}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{float}
\usepackage[usenames]{xcolor}

@@

What follows is an explanation of the initial stages of the algorithm on a very
simple case.  The diagram below shows a 64- base-pair ``genome'' that has had
three 32-base-pair error-free reads at a constant separation of 12 base pairs.
Note that the genome is two-stranded.  {\tt A} always pairs with {\tt T} and
{\tt C} always pairs with {\tt G}, but the two strands run in opposite
directions.  Reads 1 and 2 come from the top (reverse strand), while read 3
comes from the bottom (forward strand).  The reads are directed as shown (e.g.
Read 1 begins with {\tt ACAGC}, reading from right to left, not {\tt TGAGC}).

%make out.{bi,}digraph.{pdf,dot} out.reduced.{bi,}digraph.pdf
%out.reduced.collapsed.{bi,}digraph.pdf SAMPLE_SIZE=64 USE_PIRS=false
%MIN_OVERLAP_LEN=8 READ_LEN=32 READ_SEP=12 -B
\vspace{0.5cm}
\nopagebreak

@@

From looking at the diagram, it is apparent that Read 1 overlaps with Read 2 by
20 base pairs, Read 1 overlaps with Read 2 by 20 base pairs, and Read 1 overlaps
with Read 3 by 8 base pairs.  These overlaps are computed by the {\tt
compute-overlaps} program.  Note that an overlap may match either the forward or
reverse-complement sequence.

Before building the string graph, {\tt remove-contained-reads} is run to remove reads
that are fully contained by another read.  There are no such reads in this
example.

Given the set of overlaps, the next step is to build a bidirected or directed
string graph.  Both these graphs are shown on the next page.  In the directed
graphs, a node $n.B$ means the beginning of read $n$, while the node $n.E$ means
the end of read $n$.

Consider graph (2A), the directed string graph after transitive reduction, and
consider the walk $1.B \to 2.B \to 3.E$.  This represents a way in which reads
1, 2, and 3 assemble together consistently.  In the simplest case, all the nodes
in a path would be $E$ nodes, and this would correspond to a sequence of reads
that assemble together as they are, with no reverse-complementation.  However,
in this case, we walk from the {\it beginning} of read 1 to the {\it beginning}
of read 2, indicating that they are laid out in reverse-complement order, before
reaching the {\it end} of read 3, which is in forward order.  This
reconstruction is consistent with how the reads were actually sampled from the
genome (shown earlier).

The other component of graph (2A) produces an equivalent assembly, but it lays
out the reads in the opposite order and assembles the other strand of the
genome.

Graph (2B) is equivalent to (2A), but combines every node $n.B$ and $n.E$ into
one node $n$ to deal with the double-strandedness in a more concise manner.  The
edges are bidirected.  A bidirected edge may be traversed either way in a walk,
but if a vertex is entered through an inward head, it must be left through an
outward head, and vice versa.

In both the directed and bidirected graphs, the edges are labeled by DNA
sequences, although in the images, only the lengths of the sequences are shown.

\newpage
\begin{tabular}{|p{0.5\textwidth}|p{0.5\textwidth}|} \hline
\includegraphics[scale=0.6]{out.digraph-crop.pdf}
\newline 1A. Directed string graph of original data
&
\includegraphics[scale=0.6]{out.bidigraph-crop.pdf}
\newline 1B. Equivalent bidirected string graph to (1A)
\\ \hline

\includegraphics[scale=0.6]{out.reduced.digraph-crop.pdf}
\newline 2A. Graph (1A) after transitive reduction.  Read 2 is between reads 1
and 3, so there is no need for edges directly between 1 and 3.
&
\includegraphics[scale=0.6]{out.reduced.bidigraph-crop.pdf}
\newline 2B. Graph (1B) after transitive reduction, or the equivalent bidirected
string graph to (2A)

\\ \hline

\includegraphics[scale=0.6]{out.reduced.collapsed.digraph-crop.pdf}
\newline 3A. Graph (2A) after collapsing unbranched paths
&
\includegraphics[scale=0.6]{out.reduced.collapsed.bidigraph-crop.pdf}
\newline 3B. Graph (2B) after collapsing unbranched paths, or the equivalent
bidirected string graph to (3A).

\\ \hline

\end{tabular}
\newpage

%\begin{minipage}[c]{0.4\linewidth}
%\includegraphics[scale=0.6]{out.digraph-crop.pdf}
%\captionof{figure}{Directed string graph of original data}
%\end{minipage}
%\begin{minipage}[c]{0.4\linewidth}
%\includegraphics[scale=0.6]{out.bidigraph-crop.pdf}
%\captionof{figure}{Equivalent bidirected string graph of original data}
%\end{minipage}

%\hrule

%\begin{minipage}{0.5\linewidth}
%\includegraphics[scale=0.6]{out.reduced.digraph-crop.pdf}
%\end{minipage}
%\begin{minipage}{0.5\linewidth}
%\includegraphics[scale=0.6]{out.reduced.bidigraph-crop.pdf}
%\end{minipage}

%\hrule

%\begin{minipage}{0.5\linewidth}
%\includegraphics[scale=0.6]{out.reduced.collapsed.digraph-crop.pdf}
%\end{minipage}
%\begin{minipage}{0.5\linewidth}
%\includegraphics[scale=0.6]{out.reduced.collapsed.bidigraph-crop.pdf}
%\end{minipage}
